---
title: "Data Analytics and Machine Learning PS3"
author: "Huanyu Liu, Yong Jia Tan, Tongsu Peng, Sejal Bharati"
date: "4/22/2019"
output: pdf_document
---
# Question 1
a.  

```{r, echo=TRUE}
library(foreign)
library(data.table)
df = read.dta('LendingClub_LoanStats3a_v12.dta')
# (i)
df = df[df$loan_status == 'Fully Paid' | df$loan_status == 'Charged Off',]
df$default = 0
df[df$loan_status == 'Charged Off', 'default'] = 1
# (ii)
num_defaults = sum(df$default)
default_rate = num_defaults / nrow(df)
print(paste('average default rate in the sample is',default_rate))
```  
b.  
(i)
```{r, echo=TRUE}
out = glm(formula = default ~ grade, family = 'binomial', data = df)
summary(out)
```   
The larger the coefficients are, the higher probability there would be default.  
  
(ii)  
```{r, echo=TRUE}
# (ii)
test_stat = out$null.deviance - out$deviance
k = out$df.null - out$df.residual
pval_chisq = 1 - pchisq(test_stat,df = k)
print(pval_chisq)
```  
The p-value of chi-square test is 0, so that we can reject the null hypothesis, which mean at least one coefficient should not be 0. Therefore, this model performs better than the null model.  
  
(iii)  
```{r}
# (iii)
phat = predict(out, type = 'response')
ranks = rank(phat, ties.method = 'first')
deciles = cut(ranks, quantile(ranks, probs=0:10/10), include.lowest=T)
deciles = as.numeric(deciles)
df2 = data.frame(deciles=deciles,phat=phat,default=df$default)
lift=aggregate(df2,by=list(deciles),FUN="mean",data=df2)
lift=lift[,c(2,4)]
lift[,3]=lift[,2]/mean(df$default)
names(lift)=c("decile","Mean Response","Lift Factor")
lift

simple_roc <- function(labels, scores){
  labels <- labels[order(scores, decreasing=TRUE)]
  data.frame(TPR=cumsum(labels)/sum(labels), FPR=cumsum(!labels)/sum(!labels), labels)
}
roc = simple_roc(df$default == '1', phat)
TPR = roc$TPR
FPR = roc$FPR
library(ggplot2)
q = qplot(FPR,TPR,xlab = 'FPR', ylab = 'TPR', col = I('red'), main = 'ROC Curve, Logistic Default Model', size = I(0.5))
q = q + geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1)) + theme_bw()
q
```  
The mean response of lift table should be higher with higher decile, same with lift factor.  
The x-axis is the false positive rate, and the y-axis is the true positive rate. The line of ROC curve is tracing out the true and false positives for different cutoffs.  
According to the result lift table and ROC, it is better than random guess.  
  
(iv)  
```{r}
#(iv)
profits = sum(df$default == '0')*(1-FPR)*1
loss = sum(df$default == '1')*(1-TPR)*10
net = profits - loss
index = which.max(net)
cutoff = phat[index]
q + geom_point(aes(x = FPR[index], y = TPR[index], col = I('green'), size = I(2)))
cut_off_p = TPR[index]
print(paste('The cutoff probability should be',cut_off_p))
```
c. 
(i)  
```{r}
#(i)
out2 = glm(formula = default ~ loan_amnt + annual_inc, family = 'binomial', data = df)
summary(out2)
phat2 = predict(out2, type = 'response')
deciles2 = cut(phat2, quantile(phat2, probs=0:10/10), include.lowest=T)
deciles2 = as.numeric(deciles2)
df3 = data.frame(deciles=deciles2,phat=phat2,default=df$default)
lift2=aggregate(df3,by=list(deciles2),FUN="mean",data=df2)
lift2=lift2[,c(2,4)]
lift2[,3]=lift2[,2]/mean(df$default)
names(lift2)=c("decile","Mean Response","Lift Factor")
lift2
roc2 = simple_roc(df$default == '1', phat2)
TPR2 = roc2$TPR
FPR2 = roc2$FPR
roc <- cbind(roc,Model = "V1")
roc2 <- cbind(roc2, Model = "V2")

New_ROC <- rbind(roc, roc2)
q = qplot(FPR,TPR, data = New_ROC, xlab = 'FPR', ylab = 'TPR', col = Model, main = 'ROC Curve, Logistic Default Model 2', size = I(0.5))
q + geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1), col = I('black')) + theme_bw()
```
According to the ROC curve, the area under grade model is larger than that of the alternative model, so the grade model is better.
  
(ii)  
```{r}
#(ii)
out3 = glm(formula = default ~ loan_amnt + annual_inc + term + int_rate, family = 'binomial', data = df)
summary(out3)
phat3 = predict(out3, type = 'response')
deciles3 = cut(phat3, quantile(phat3, probs=0:10/10), include.lowest=T)
deciles3 = as.numeric(deciles3)
df4 = data.frame(deciles=deciles3,phat=phat3,default=df$default)
lift3=aggregate(df4,by=list(deciles3),FUN="mean",data=df4)
lift3=lift3[,c(2,4)]
lift3[,3]=lift3[,2]/mean(df$default)
names(lift3)=c("decile","Mean Response","Lift Factor")
lift3
roc3 = simple_roc(df$default == '1', phat3)
TPR3 = roc3$TPR
FPR3 = roc3$FPR
roc3 <- cbind(roc3, Model = "V3")

New_ROC <- rbind(New_ROC, roc3)
q = qplot(FPR,TPR, data = New_ROC, xlab = 'FPR', ylab = 'TPR', col = Model, main = 'ROC Curve, Logistic Default Model 2', size = I(0.5))
q + geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1), col = I('black')) + theme_bw()
```  
  
R treats term-variable as dummy variable.  
The larger the coefficient means that variable affects more on the default probability. Positive coefficients means the smaller the variable the lower default probability, and the opposite way for the negative coefficients.  
According to the ROC curve, this new model is better.  
The default rate may be more relevant to the characteristics of the specific loan than to the grade of the borrowers.  
  
(iii)  
```{r}
#(iii)
df$int_rate_sq = df$int_rate ^ 2
out4 = glm(formula = default ~ loan_amnt + annual_inc + term + int_rate + int_rate_sq, family = 'binomial', data = df)
summary(out4)

```
Yes, the coefficent on the square of interest rate is significant.  
The positive coefficient on interest rate means that default rate will be lower with lower interest rate. While the negative coefficient on interest rate means that default rate will be higher with lower square of interest rate.  
Because the square of interest rate is too small, overall interest rate will dominate its square. Therefore, default rate will be lower with lower interest rate in general.
