---
title: "Project3"
author: "Huanyu Liu"
date: "1/30/2019"
output: pdf_document
---

# Problem 1

$$\begin{aligned}
P(Y_2 > 5) &= 0.974\\
E(X_2^{1/3}) &= 0.658903\\
E(Y_3) &= 25.8096\\
E(X_2Y_2 1(X_2 > 1)) &= 4.09305\\
\end{aligned}$$  

#Problem 2
$$\begin{aligned}
E((1+X_3)^{1/3}) = 1.32474\\
E((1+Y_3)^{1/3}) = 1.33535\\
\end{aligned}$$

# Problem 3
(a)  
```{r,eval=TRUE}
library("knitr")
df = read.csv("Project3_P3.csv")
kable(df,caption="Call Option stas")
```
  
  
```{r,eval=TRUE}

greeks = c("Delta", "Gamma", "Theta", "Vega", "Rho")
for (i in c(1:5)){
  plot(x = df[[1]], y = df[[i+3]], xlab = "S0", ylab = greeks[i], type = "b", pch = 18)
}
```
  
# Problem 4
T = 0.5  
K = 50  
Full Truncation: C1 = 2.59491    
Partial Truncation: C2 = 2.59491  
Reflection Method: C3 = 2.59491  

# Problem 5
(d)
```{r, eval=TRUE}
df2 = read.csv("Project3_P5.csv", header = FALSE)
for (i in c(1,3,5)){
  plot(x = df2[[i]], y = df2[[i+1]], xlab = "x", ylab = "y",pch = 20)
}
```

Both Uniform and 2-dimension Halton with base 2 and 7 can produce random sequence numbers, however, 2-dimension Halton with base 2 and 4 is not a sequence of random numbers. There is correlation of the sequences by Halton with base 2 and with base 4.  
  
  
    
      
      
(e)
$$\int_0^1 \int_0^1 e^{-xy}(sin6\pi x + cos^{\frac{1}{3}}2\pi y)dxdy$$
By (2,4), value = -0.0048839  
By (2,7), value = 0.0261144  
By (5,7), value = 0.0261637